Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
- data aggregation?

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

qplot(x = age, y = friend_count, data = pf)

# the same like above:
qplot(age, friend_count, data = pf)

# get some boundaries:
summary(pf$age)

# the same with ggplot syntax:
ggplot(aes(x = age, y = friend_count), data = pf) + geom_point() +
  xlim(13, 90)


```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

```

***

### Overplotting
Notes:

```{r Overplotting}
# overplotting?
# has to be at leat 20 to get dark dot!
ggplot(aes(x = age, y = friend_count), data = pf) + 
  # geom_point(alpha = 1/20) +
  geom_jitter(alpha = 1/20) + 
  xlim(13, 90) 

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:
coord trans add a layer to the plot that transforms friend_count using the square root function:
```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13, 90) + 
  coord_trans(y = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}

```

***

### Overplotting and Domain Knowledge
Notes:

***
some common dyplr functions:
filter()
group_by()
mutate()
arrange()


### Conditional Means
Notes:

```{r Conditional Means}
install.packages('dplyr')
library(dplyr)

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups, 
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) # n can be only used for summarise
pf.fc_by_age <- arrange(pf.fc_by_age, age)


head(pf.fc_by_age)

# other option:

pf.fc_by_age <- pf %>% ## chain !
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)


head(pf.fc_by_age)  


```


Create your plot!

```{r Conditional Means Plot}
# plot fc vs age:
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) + 
  geom_line() +
  xlim(13, 90) 



```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}

ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, 
             position = position_jitter(h = 0),
             color = 'orange') +
  xlim(13, 90) + 
  coord_trans(y = "sqrt") +
  geom_line(stat = 'summary', fun.y = mean) + 
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .1),
            linetype = 2,
            color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .9),
            linetype = 2,
            color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .5),
            color = 'brown') 

# cartesian coordinates:

ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, 
             position = position_jitter(h = 0),
             color = 'orange') +
  coord_cartesian(xlim = c(13, 70), ylim = c(0, 1000 )) +
  geom_line(stat = 'summary', fun.y = mean) + 
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .1),
            linetype = 2,
            color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .9),
            linetype = 2,
            color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, 
            fun.args = list(probs = .5),
            color = 'brown') 


```

#### What are some of your observations of the plot?
Response:
-- almost noone has more than 2k users
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
#with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
# install.packages('alr3')
# library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

